---
title: "changepoint-forecast"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{changepoint-forecast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(changepoint.forecast)
```

`changepoint.forecast` is an R package for detecting changepoints using forecast errors. This package is currently in development and thorough testing has not been performed.

The main function is `cptForecast`. This function takes forecast errors along with additional information such as the length of the training period where the errors are assumed stable and the type of CUSUM detector to use. The function works by running the CUSUM detector on the forecast errors and the squared forecast errors (the former is designed to detect mean changes and the later for second-order changes). The theoretical critical values of the CUSUM detectors can either be calculated with `critValue='Simulate'` and specifying additional arguments or can be looked up from a pre-calculated table of values using `critValue='Lookup'`. Otherwise a manual critValue can be entered. The type-1 error rate `alpha` can be set using the argument `alpha`.

The returned object is an S4 class `cptFor`. This object contains the time when a changepoint was detected `tau` for the forecast errors and `tau2` for the squared forecast errors, along with other additional information that can be found in the documentation. Slots of this class can be accessed using `@` as shown below and a plot methods is also available for this class. 


The example below simulates 500 forecast errors, the training period will be set to `m=300` and at time point 400 there is a changepoint in the mean. 

```{r basicExample}
set.seed(1)
errors = c(rnorm(400),rnorm(100,2))
m = 300

out = cptForecast(errors=errors, m=m, detector='PageCUSUM', critValue='Lookup', alpha=0.05)
out@tau
out@tau2
plot(out)

```





